---
title: "Project Title Here"
author: "Author Name Here"
date: "Current Date Here"
---















#### ----------------------------------


### **Module 8 (M8) - Sports II**

*Purpose: What is Statistics/Data Science (DS) Question we want to answer?*


3. **Touchdown to Interception Ratio**: Which quarterbacks have a touchdown to interception ratio greater than 2?
- *Manipulation*: Calculate the touchdown to interception ratio for each quarterback and filter those with a ratio greater than 2.
- *Visualization*: Create a scatter plot with touchdowns on the x-axis and interceptions on the y-axis, highlighting quarterbacks who meet the criteria.




##### The Extraction

```{r}
library(tidyverse) ## ggplot2 and dplyr
library(gato365dsh2024)
```

##### The Tranformation

```{r}
td_per_int_passing_df <- passing_stats_df %>% 
  distinct(.,.keep_all = TRUE) %>% 
  filter(year == 2023) %>% 
  select(player,int,td)  %>% 
  mutate(td_per_int = ifelse(int == 0, NA, td/int)) %>% 
  filter(td_per_int <=4 & td_per_int >=1) %>% 
  arrange(td_per_int) %>% 
  mutate(player = factor(player, levels = unique(player)))
```

##### The Visualization


```{r}
td_per_int_passing_df  %>% 
  ggplot(aes(x=player, y = td_per_int)) +
  geom_point() + 
  geom_text(aes(label=round(td_per_int, 2)), hjust=-0.1, vjust=-0.5) +
  labs(x = "Player", y = "TD/INT Ratio", title = "TD/INT Ratio for Each Player in 2023") +
  coord_flip() +
  theme_custom()
```















## Weekend 3 Outline


More questions



##### The Extraction

##### The Tranformation

##### The Visualization


- Hackathon




