

```{r}
library(gato365dsh2024)
```


## Weekend 1 Outline


<!-- SECTION: Module 0 (M0) -->
### **Module 0 (M0)**
- Introduction and Overview


#### Overview of the Hackathon Process

The Data Science Hackathon is designed to engage high school students in the practical aspects of data science through a focused competition. Over the course of two weekends, participants will undergo a learning phase through instructional videos and can get help from mentors, followed by a weekend dedicated to applying their new skills in a hackathon format. This process aims to not only introduce students to the field of data science but also to challenge them to solve real-world problems using data from interesting domains like Spotify and sports.

#### Core Objectives of the Hackathon

1. **Igniting a Passion for Data Science: ** Launch high school students on an exhilarating journey into the world of programming and data science, setting the stage for future innovations and discoveries.

2. **Hands-On Experience with Authentic Data: ** Dive deep into the vast ocean of real-world data, where students will learn to navigate, analyze, and extract valuable insights from genuine datasets from intriguing domains like Spotify and sports.

3. **Collaboration and Mentorship: ** Foster a dynamic environment where students engage in an interactive project alongside mentors from Cal Poly. This collaboration is designed to bridge theoretical knowledge with practical application, encouraging students to solve complex problems through teamwork and expert guidance.

#### Foundational Skills in Programming and Data Science

Participants will gain essential skills in:
- Basic programming concepts and syntax, likely using R.
- Data manipulation and analysis using packages like dplyr and stringr (for R).
- Data visualization techniques using ggplot2 (for R).
- Understanding of how to approach data science problems from hypothesis to insight.

#### Introduction to Posit Cloud and Its Role in the Hackathon

Posit Cloud serves as the primary platform for the hackathon, providing an accessible, web-based environment for coding. It offers:
- An integrated development environment (IDE) for R and Python, removing the need for local software installation.
- Easy access to project workspaces, facilitating seamless collaboration among participants.
- A platform for mentors to share code snippets, instructional materials, and feedback in real-time.

#### How to Use Posit

Participants will learn to:
- Navigate the Posit Cloud interface to create and manage projects.
- Utilize the IDE for writing and executing code.
- Share projects with mentors for review and collaboration.
- Deploy applications or visualizations directly from Posit Cloud.

#### Mentorship and Support Available

Mentorship plays a crucial role in the hackathon, with support provided by:
- College students and professionals in data science, offering guidance through office hours and discussion forums.
- Structured feedback sessions, where participants can present their progress and receive constructive advice.
- Access to a repository of resources, including tutorials, datasets, and examples specifically curated for the hackathon challenges.

#### Quiz Engagement

- Complementing each video, you'll find a quiz designed to sharpen your newly acquired skills.
- Enjoy the freedom of unlimited attempts to ensure mastery of the material.
- Don't hesitate to reach out to your mentors for assistance and insights.













<!-- SECTION: Module 1 (M1) -->



### **Module 1 (M1)**

**How does this tie to the journey?**
Version 1:
These topics tie into the data science learning journey by providing the foundational tools and frameworks necessary for data manipulation, analysis, and visualization. Understanding how to install and utilize R packages enables learners to efficiently process and analyze large datasets, apply statistical models, and generate insightful visualizations, which are critical skills in the field of data science.

Version 2:

"Hey there! So, these topics we're talking about are super important if you're interested in data science—that's like being a detective, but for numbers and patterns. Learning how to use R, which is a special tool for data science, is like learning how to use a new app on your phone. It helps you sort through tons of information, find cool trends, and make those awesome graphs and charts you see in science fairs. It's all about making sense of the data and telling a story with it. Pretty neat, right?" 😄📊



#### **Part 1 (P1)** Basics of what is R

##### Concept 1: R and R Studio

- What is R?
- Solution: R is a programming language and environment designed for     
statistical computing and graphics.

- What is R Studio?
- Solution: RStudio is an integrated development environment (IDE) that 
provides a user-friendly interface for working with R.  

- Why R Studio?
- Solution: RStudio enhances the R experience by offering a powerful, 
cohesive toolset for data analysis, visualization, and programming in one 
interface.  

##### Concept 2: Panes 

- **Console**: The Console in R is the interactive interface allowing users to execute R commands and view their results immediately.

- **Files (Types)**:
- **R File**: An R file (`.R`) contains scripts written in R language, used for data analysis and visualization.
- **Quarto File**: A Quarto file (`.qmd`) is used for creating dynamic documents that can combine code execution with narrative text, images, and other content.

- **Environment**: The Environment in R refers to the workspace that holds user-defined variables, data sets, and functions during an R session.



##### Concept 3: R Packages and Libraries

- **What is an R package and library?**  
An R package is a collection of functions, data, and compiled code in a well-defined format, while the library is a directory where the packages are stored and managed.

- **How to install and load R packages?**  
To install R packages, use the `install.packages("packageName")` command, and to load them into an R session, use `library(packageName)`.

- **Some commonly used R packages**  
Commonly used R packages include `ggplot2` for data visualization, `dplyr` for data manipulation, `tidyr` for data tidying, and `shiny` for interactive web applications.




#### ----------------------------------




#### **Part 2 (P2)** Data Types - Vectors and Data frames

**How does this tie to the journey?**
**"Imagine organizing your game inventory: Vectors in R are like backpack slots, holding similar items, while data frames are like chests for different items, aiding data scientists in managing vast info like gamers strategize to win!" 🎮📊**

To create a new R code chunk in RStudio, you can use the shortcut:
- **Mac**: Cmd + Option + I
- **Windows/Linux**: Ctrl + Alt + I

#### Concept 1: Variable Type

- **Categorical Variables**: Represent categories and can have more than two possible values, like species types in the iris dataset.
```{r}
cat_var <- iris$Species # Species in the iris dataset
```

- **Categorical (Binary) Variable**: A special case of categorical variables with two possible values, such as "Yes" or "No".
```{r}
binary_var <- c("Yes", "No", "Yes") # Examples of binary variables
```


- **Quantitative Variables**: Represent numeric values at which numeric operations make sense, like Sepal Length in the iris dataset.
```{r}
cat_var <- iris$Sepal.Length # Species in the iris dataset
```


#### Concept 2: Vectors

- **Introducing Vectors**: Vectors are the simplest way to store data in R, holding elements of the same type.
```{r}
num_vec <- c(1, 2, 3) # Numeric vector
char_vec <- c("apple", "banana", "cherry") # Character vector
```

- **Numeric Vector**: Contains numeric (quantitative) values.
```{r}
num_vec <- c(1.1, 2.2, 3.3) # Example of a numeric vector
```

- **Character Vector**: Holds text strings. In R, strings and characters are treated similarly, used for storing textual data.
```{r}
char_vec <- c("apple", "banana", "cherry") # Example of a character vector
```

- **Accessing Vector Elements**: Elements within a vector can be accessed using square brackets `[]` and the index of the element.
```{r}
char_vec[2] # Access the second element of char_vec
```





#### Concept 3: Data Frames

- **What is a data frame?**
A data frame is a table or a two-dimensional array-like structure in which each column contains values of one variable and each row contains one set of values from each column.
```{r}
df <- data.frame(age = c(15,16,17), grade_in_class = c("A", "B", "A-"))
```

- **What are rows?**
Rows in a data frame represent observations or records.
```{r}
df[1, ] # Access the first row
```

- **What are columns?**
Columns in a data frame represent variables.
```{r}
df[, 1] # Access the "Number" column
```

- **How to access rows and columns within a data frame?**
You can access rows and columns within a data frame using square brackets `[,]` with row indices before the comma and column indices or names after the comma.
```{r}
df[1, 2] # Access the first row and "Number" column
```





#### Concept 4: Lists

- **What is a list?**
A list in R is an ordered collection of objects, which can be of different types and structures.
```{r}
lst <- list(number = 42, name = "R", data = df)
```

- **How to access items within a list**
You can access items within a list using double square brackets `[[]]` for index or single square brackets `[]` with the item name.
```{r}
lst[["name"]] # Access the "name" item
lst$name # Another way to access the "name" item
lst[[1]]
```








<!-- SECTION: Module 2 (M2) -->



### **Module 2 (M2)**

#### **Part 1 (P1)** Extraction and Transformation (ET) - Box Office Analysis

---------------------------------------
**How does this tie to the journey?**
Envision yourself as a data analyst at a bustling movie studio. You're tasked with identifying blockbuster trends that capture audience interests. You'll harness the power of data to uncover hidden patterns and drive successful film projects. For example investigating age and genres of movies.
---------------------------------------

##### Concept 1: Extract - Read in CSV

(GO TO EXCEL)

To start analyzing Excel data in R, we first need to read the data into R using the `readxl` package's `read_xlsx()` function.

```{r}
library(readr)
movies_df <- read_csv("IMDB_Genres.csv")
```

##### Concept 2a: Transform - Select Variables

The `%>%` symbol in R is called the pipe operator, and it's used to pass the output of one function directly into the next, making your code cleaner and easier to read. It's like saying "then do this" between steps in your data analysis process.

The `select` function in R, part of the `dplyr` package, is used to choose specific columns from a dataset by their names or positions, which can also include renaming or dropping columns. It's like picking your favorite songs to create a custom playlist from your entire music library.

```{r}
library(dplyr)
title_year_genre_df <- movies_df %>%
  select(Title, Year, Genre)
```


##### Concept 2b: Transform - Mutate Variables

The mutate function in R is used to add new variables to a dataset or change existing ones. It’s like updating your playlist by adding new songs or changing the order of the tracks.

```{r}
age_df <- title_year_genre_df %>%
  mutate(Age = 2023 - Year)
```



##### Concept 3: Visualize - Histogram

Think of a histogram like a bar chart that shows you how many times something happens. It's great for seeing how often certain numbers come up.

```{r}
library(ggplot2)

age_df %>% 
  ggplot(aes(x=Age)) + 
  geom_histogram(binwidth = 10)
```

- **What type of variable can be used within Histogram?**
- You can use a numeric variable, like the Age of movies.
- **Change Bin Size**
- The `binwidth` parameter lets you decide how wide each bar should be. Smaller `binwidth` means more bars!

##### Concept 4: Visualize - Bar Plot
Bar plots are like histograms, but they're used for categories, like movie genres.

```{r}
age_df %>% 
  ggplot(aes(x=Genre)) + 
  geom_bar()
```

- **What type of variable can be used within bar plot?**
- Categorical variables, like 'Comedy' or 'Action', are perfect for bar plots. They show you how many movies are in each genre.








#### ----------------------------------



#### **Part 2 (P2)** Visualization (V) - Marketing Insights from Viewership Data
---------------------------------------
**How does this tie to the journey?**
Now, imagine fine-tuning the studio's marketing strategy. Through data, you dissect viewer preferences, pinpointing exactly what pulls audiences into theaters. Your insights fuel targeted campaigns that resonate deeply and fill seats.

---------------------------------------


##### Concept 1: Transform - Filter Variables

Lets Learn about filter function.  

The `filter` function in R is used to subset a data frame, keeping only the rows that meet certain conditions. It's like a gatekeeper that only lets through the data that matches your specified rules.

```{r}
# Filter a quantitative variable (e.g., Rating greater than 8)
highly_rated_movies <- movies_data %>% filter(Rating > 8)

# Filter a categorical binary variable (e.g., Movies that are Drama)
drama_movies <- movies_data %>% filter(is_drama == "Drama")

# Filter a categorical variable (e.g., Movies of the 'Action' genre)
action_movies <- movies_data %>% filter(Genre == "Action")
```

##### Concept 2: Transform - Mutate Variables

Let's talk about mutate again.
```{r}
# Change the levels name (e.g., Rename 'Not Drama' to 'Other')
movies_df <- movies_df %>% mutate(is_drama = recode(is_drama, 'Not Drama' = 'Other'))

# Turn a quantitative variable into a Boolean variable (e.g., High Revenue)
movies_df <- movies_df %>% mutate(high_revenue = Revenue > mean(Revenue, na.rm = TRUE))
```

##### Concept 3: Visualize - Boxplot
- **What type of variable can be used within box plot?**
- Numeric variables, such as 'Rating' or 'Runtime', are used in box plots to show the distribution of data.

```{r}

movies_df %>% 
  ggplot(aes(y=Rating)) + 
  geom_boxplot()
```

This code will produce a boxplot showing the spread of movie ratings in your dataset.

##### Concept 4: Visualize - Boxplot across multiple groups
- **What type of variable can be used within box plot with multiple groups?**
- A categorical variable, like 'Genre', can be used to compare the numeric distribution across different groups.


To compare the distribution of a numeric variable across different categories, like 'is_thriller', you can use this R code:

```{r}
movies_df %>% 
  ggplot(aes(x=is_thriller, y=Rating)) + 
  geom_boxplot()
```

This will create a boxplot for each is_thriller, allowing you to compare the ratings across thriller and non thriller movies.



<!-- SECTION: Module 3 (M3) -->





### **Module 3 (M3)**

#### **Part 1 (P1)** ET - Community Insights for Public Policy

---------------------------------------

**How does this tie to the journey?**

Picture yourself collaborating with state officials, wielding data to inform policy decisions. Your analysis of community demographics and needs illuminates the path to impactful legislation and improved civic life.
---------------------------------------
##### Concept 1: Extract - htmltab



Extracting data is our first step. We'll use a function similar to `htmltab` to pull the list of counties directly from the Wikipedia page.

The `which` argument lets us specify which table on the page we're interested in, as pages often contain multiple tables.

```{r}
# library(htmltab)
# library(dplyr)
url <- "https://en.wikipedia.org/wiki/List_of_counties_in_California"
california_counties_df <- htmltab(url, which = 2)


california_counties_df <- california_counties_df %>% 
  rename(Population = `Population(2023)`,
         Established = Est.)

```


##### Concept 2a: Transform - str_remove_all & as.numeric

First, we remove commas from the population figures, making them easier to work with as numeric values. 

Then, we convert the cleaned population data from strings to actual numeric values.

```{r}
# library(stringr)
## I am going to keep using clean clean_california_counties_df
## This is not always done in the real world
clean_california_counties_df <- california_counties_df %>%
  mutate(Population = str_remove_all(Population,",")) %>% 
  mutate(Population = as.numeric(Population))

```

##### Concept 2b: Transform - Adding & Select a variables

```{r}

## To create more variables
clean_california_counties_df <- clean_california_counties_df %>%
  mutate(State = "CA") %>% 
  mutate(Established = as.numeric(Established)) %>% 
  mutate(Est_before_1853 = ifelse(Established<=1853,"Yes","No"))

## To make the code more neat
clean_california_counties_df <- clean_california_counties_df %>% 
  select(State, Population, County, Est_before_1853)

```



##### Concept 3: Visualize - geom violin plot

A violin plot is a method of plotting numeric data and can be thought of as a combination of a box plot and a kernel density plot. It provides a deeper understanding of the distribution of the data.

```{r}

library(ggplot2)
clean_california_counties_df %>% 
  ggplot(aes(x=Est_before_1853, y=Population)) + 
  geom_violin(trim=FALSE) 
```

This code will create a violin plot for the population of counties in California, giving a visual representation of the distribution of population sizes.

##### Concept 4: Visualize - labels & theme_custom

Adding labels and themes can make your plots more informative and aesthetically pleasing. The `theme_custom()` function gives a clean, white background to your plots.

```{r}
# library(scales)
# library(gato365dsh2024)
clean_california_counties_df %>% 
  ggplot(aes(x=Est_before_1853, y=Population)) + 
  geom_violin(trim=FALSE) +
  labs(title="Population Distribution in California Counties \n Based on when they were established",
       x="Established Before 1853",
       y="Population") #+
# scale_y_continuous(labels = label_number(accuracy = 0.1)) +
# theme_custom()
```

This code will produce a violin plot with a title and axis labels, set against a clean, white background theme. It's a great way to present your data in a clear and professional manner.


#### ----------------------------------





#### **Part 2 (P2)** TV - Enhancing State Services with Data
---------------------------------------

**How does this tie to the journey?**
As you dive into the world of social media analytics, your focus shifts to understanding the digital landscape. By examining the data from the most-followed TikTok accounts, you develop insights into online engagement trends. This knowledge helps you devise innovative approaches to digital marketing, ultimately enhancing the online presence and influence of your state's initiatives.
---------------------------------------


##### Concept 1: Extract the data


```{r}
library(htmltab)
url <- "https://en.wikipedia.org/wiki/List_of_most-followed_TikTok_accounts"
tiktok_df <- htmltab(url, which = 1)

```

##### Concept 2: Transform - mutate Followers and Likes into numeric

```{r}
## Again we 
clean_tiktok_df <- tiktok_df %>%
  mutate(Followers_Millions = as.numeric(`Followers(millions)`)) %>%
  mutate(Likes_Billions = as.numeric(`Likes(billions)`))
```



##### Concept 3: Transform - str_detect: based on description detect if person is social media personality or not  
```{r}
clean_tiktok_df <- clean_tiktok_df %>%
  mutate(Is_Social_Media_Personality = ifelse(str_detect(Description, "Social media personality"), TRUE, FALSE))
```





##### Concept 4: Visualize - geom point: follower vs likes

```{r}
## 
clean_tiktok_df %>% 
  ggplot(aes(x=Followers_Millions, y=Likes_Billions)) +
  geom_point() + 
  theme_custom()
```




##### Concept 5: Visualize - color by social media personality or not and labels 
```{r}
clean_tiktok_df %>% 
  ggplot(aes(x=Followers_Millions, y=Likes_Billions, color=Is_Social_Media_Personality)) +
  geom_point() +
  # scale_color_manual(values=c("blue", "red")) +
  labs(title="Followers vs Likes on TikTok",
       x="Number of Followers",
       y="Number of Likes",
       color="Is Social Media Personality") +
  theme_custom()

```








<!-- SECTION: Module 4 (M4) -->




### **Module 4 (M4)**

#### **Part 1 (P1)** ET - Gaming Market Dynamics
---------------------------------------
**How does this tie to the journey?**
Step into the virtual realm of a game developer. With data as your guide, explore the ebb and flow of gaming trends. Use this insight to innovate, crafting the next gaming sensation that captivates a global audience.
---------------------------------------


##### Concept 1: Extract Game Data with `htmltab`

```{r}
library(htmltab)
games_2023_df <- htmltab("https://en.wikipedia.org/wiki/2023_in_video_games", which = 4)
```
**Explanation**: `htmltab()` is used to extract data from HTML tables on web pages. Here, it retrieves the fourth table from the specified Wikipedia page, containing information about video games released in 2023.

##### Concept 2: Transform Data with `dplyr`

```{r}
library(lubridate)
library(dplyr)
fix_time_games_2023_df <- games_2023_df %>% 
  mutate(Release = mdy(Release)) %>%  # Convert release date to Date format
  mutate(`Average score` = as.numeric(`Average score`)) %>%  # Convert average score to numeric
  rename(Average_Score = `Average score`, Platforms = `Platform(s)`)  # Rename columns for easier reference
```
**Explanation**:
- `mutate(Release = mdy(Release))`: Uses `lubridate`'s `mdy()` function to convert the "Release" column from a string to a Date object.
- `mutate(Average score = as.numeric(Average score))`: Converts the "Average score" column to numeric type for calculations.
- `rename()`: Renames columns to remove spaces and special characters, facilitating easier access and manipulation.

##### Concept 3: Group by & Summarise

```{r}
# summary_df <- 
fix_time_games_2023_df %>% 
  group_by(Platforms) %>% 
  summarise(mean = mean(Average_Score, na.rm = TRUE),  # Calculate mean average score
            sd = sd(Average_Score, na.rm = TRUE),  # Calculate standard deviation
            ss = n())  # Count the number of entries
```
**Explanation**: Groups the data by the gaming platform, then calculates the mean score, standard deviation, and sample size for each group. This is crucial for understanding score distribution across platforms.

##### Concept 4: Visualize Data with `ggplot2` (Violin Plot)

```{r}
library(ggplot2)
library(gato365dsh2024)
# Create a violin plot
fix_time_games_2023_df %>% 
  ggplot(aes(x = Platforms, y = Average_Score)) +
  geom_violin() +
  labs(title = "Average Score by Platform", x = "Platform", y = "Average Score") +
  # ylim(88,100) +  # Set Y-axis limits
  theme_custom()
```
**Explanation**: Creates a violin plot to visualize the distribution of average scores across different gaming platforms. Violin plots are useful for showing the density of the data at different values, combined with the box plot feature.

##### Concept 5: Visualize Data - Add Mean


```{r}
fix_time_games_2023_df %>% 
  ggplot(aes(x = Platforms, y = Average_Score)) +
  geom_violin() +
  labs(title = "Average Score by Platform", x = "Platform", y = "Average Score") +
  ylim(88,100) +  # Set Y-axis limits
  # stat_summary(fun = mean, geom = "point", shape = 23, fill = "blue") + # Add mean score as a point
  theme_custom()
```



**Explanation**: Enhances the violin plot by adding a point to represent the mean average score for each platform, providing a clear visual indicator of the central tendency within the distribution. 

This breakdown transforms the script into a step-by-step data science workflow, demonstrating the journey from extracting web data to transforming and finally visualizing it in a meaningful way.

##### Concept 6: Line Plot 


```{r}
fix_time_games_2023_df %>% 
  ggplot(aes(x = Release, y = Average_Score, color = Platforms)) +
  geom_point() +
  geom_line() +
  # scale_x_date(date_breaks = "1 month", date_labels = "%b") + 
  labs(x = "Release Date", y = "Average Score", title = "Average Score by Platform over Time") +
  theme_custom()
```





#### ----------------------------------

#### **Part 2 (P2)** TV - Viral Content Creation
---------------------------------------
**How does this tie to the journey?**
Transition to a social media strategist at the forefront of digital entertainment. By understanding viral trends through data, you craft content that not only entertains but becomes a cultural touchstone, propelling your brand to new heights.
---------------------------------------
##### Concept 1: Extract: htmltabs

```{r}
games_2022_df <- htmltab("https://en.wikipedia.org/wiki/2022_in_video_games",
                         which = 5) %>% 
  mutate(Year = 2022)
games_2023_df <- htmltab("https://en.wikipedia.org/wiki/2023_in_video_games",
                         which = 4) %>% 
  mutate(Year = 2023)
```

##### Concept 2: Transform: bind_row, mutate, & rename

```{r}
games_bind_df <- games_2022_df %>% 
  bind_rows(games_2023_df) %>% 
  mutate(Release = mdy(Release)) %>% 
  mutate(`Average score` = as.numeric(`Average score`)) %>% 
  rename(Average_Score = `Average score`,
         Platforms = `Platform(s)` )
```


##### Concept 3: Visualize
```{r}
games_bind_df %>% 
  ggplot(aes(x = Release, y = Average_Score, color = Platforms)) +
  geom_point() +
  geom_line() +
  labs(x = "Release Date", y = "Average Score", title = "Average Score by Platform over Time \n By Year") +
  facet_grid(.~Year) +
  theme_custom()
```

##### Concept 4: Group by & Summarize

```{r}
games_bind_df %>% 
  group_by(Platforms,Year) %>% 
  summarise(mean = mean(Average_Score, na.rm = TRUE),
            sd = sd(Average_Score, na.rm = TRUE), 
            ss = n())
```


##### Concept 5: Visualize
```{r}
# Create a violin plot

games_bind_df %>% 
  ggplot( aes(x = Platforms, y = Average_Score)) +
  geom_violin() +
  stat_summary(fun = mean, geom = "point", shape = 23, fill = "blue") +
  labs(title = "Average Score by Platform", x = "Platform", y = "Average Score")  +
  ylim(88,100) +
  facet_grid(.~Year) +
  theme_custom()
```








<!-- SECTION: Module 5 (M5) -->


## Weekend 2 Outline





### **Module 5 (M5) - Spotify I**


#### **Part 1 (P1)** ETV - Spotify I
**How does this tie to the journey?** As a Spotify data engineer, your role is pivotal in understanding the current trends of popular artists. By analyzing data such as user listening habits, artist popularity metrics, and song play counts, you can identify patterns and trends among these artists. These insights not only help Spotify curate better playlists and make more accurate song recommendations, but they also contribute to predicting future hits. Thus, your work directly influences the user experience on Spotify, making it more personalized and enjoyable for listeners worldwide. This is how your task ties into your journey as a data engineer at Spotify.

*Purpose: What is Statistics/Data Science (DS) Question we want to answer?*
1. **Artist Album Release Patterns**: How has the number of tracks on albums released by each artist ("Dua Lipa" and "Olivia Rodrigo") changed over the years?
- *Data Manipulation*: Filter the data for the artists "Dua Lipa" and "Olivia Rodrigo". Remove duplicate tracks based on track name, then group the data by artist and album name. Count the number of tracks on each album and record the album release year.
- *Visualization*: Create a line graph showing the number of tracks on albums released per year for each artist. The lines are colored differently for each artist.


##### The Extraction
```{r}
library(gato365dsh2024)
library(dplyr)
glimpse(all_artists_df)
unique(all_artists_df$artist_name)
```

##### The Transformation
```{r}
library(ggplot2)
artist_album_df <- all_artists_df %>% 
  filter(artist_name %in% c("Dua Lipa","Olivia Rodrigo")) %>% 
  distinct(track_name, .keep_all = TRUE) %>%  # Remove duplicates based on artist and album
  group_by(artist_name, album_name) %>%  # Group by artist and album year
  summarise(count = n(), album_release_year = first(album_release_year))  # Count the number of albums
```


##### The Visualization 
```{r}
artist_album_df %>% 
  ggplot(aes(x = album_release_year, y = count, group = artist_name, color = artist_name)) +  # Plot album_year on x-axis and count on y-axis, group and color by artist
  geom_line() +  # Use line graph
  geom_point() +
  scale_color_manual(values = c("blue","green")) +
  labs(title = "Artist Album Release Patterns Over the Years", x = "Year", y = "Number of Tracks on Albums", color = "Artist Name") +  # Add labels
  theme_custom()
```














<!-- SECTION: Module 6 (M6) -->




### **Module 6 (M6) - Sports I**

*Purpose: What is Statistics/Data Science (DS) Question we want to answer?*

#### **Part 1 (P1)**: Understanding Age Distribution


**How does this tie to the journey?** 

As a data analyst for the NBA league, your role is crucial in understanding the age distribution of players. By analyzing player data, you can identify patterns and trends among different age groups. This insight not only helps in understanding the career longevity and prime playing years of players, but also aids in strategic team management and talent scouting. Thus, your work directly influences the strategic decisions made in the NBA league, contributing to the success of teams and the overall competitiveness of the league. This is how your task ties into your journey as a data analyst at the NBA.

*Purpose: What is Statistics/Data Science (DS) Question we want to answer?*

2a. **Age Distribution of Players**: How does the age distribution look across all players, and which age group is most prevalent? 

- *Data Manipulation*:  Group the players by age and count the number of players in each age group.

- *Visualization*:  Create a histogram to display the age distribution.
















##### The Extraction

```{r}
library(gato365dsh2024)
library(dplyr)
library(ggplot2)
```


##### The Transformation

```{r}
nba_players_age_df <- nba_year_2023_df %>% 
  distinct(player_name, .keep_all = TRUE) %>%  # Remove duplicates based on the player with the most games
  mutate(age_group = cut(age, breaks = seq(15, 50, 5), include.lowest = TRUE)) %>%  # Create arbitrary age groups
  group_by(age_group) %>%  # Group by age group
  summarise(freq = n()) 
```
##### The Visualization

```{r}
nba_players_age_df %>% 
  ggplot(aes(x = age_group, y = freq)) +  # Plot age_group on x-axis and frequency on y-axis
  geom_bar(stat = "identity", fill = "steelblue") +  # Use bar chart
  labs(title = "Age Distribution of NBA Players in 2023", x = "Age Group", y = "Number of Players") +  # Add labels
  theme_custom()

```




#### ----------------------------------

#### **Part 2 (P2)**: Age Distribution by Position


**How does this tie to the journey?**  
Examining age distribution by player position uncovers patterns of career duration and potential position-specific physical demands, guiding strategic decisions in team composition and player development.



```{r}
## 1b. **Age and Position Distribution of Players**: How does the age distribution look across all players conditioned on position?
nba_year_2023_df %>% 
  distinct(player_name, .keep_all = TRUE) %>%  # Remove duplicates based on the player with the most games
  mutate(position = str_split(position, "-", simplify = TRUE)[,1]) %>% 
  mutate(age_group = cut(age, breaks = seq(15, 50, 5), include.lowest = TRUE)) %>%  # Create arbitrary age groups
  group_by(position, age_group) %>%  # Group by position and age group
  summarise(freq = n()) %>%  # Count the frequency in each group
  ggplot(aes(x = age_group, y = freq, fill = position)) +  # Plot age_group on x-axis and frequency on y-axis, fill by position
  geom_bar(stat = "identity", position = "dodge") +  # Use bar chart with bars side-by-side
  labs(title = "Age Distribution of NBA Players in 2023 by Position", x = "Age Group", y = "Number of Players") +  # Add labels
  theme_custom()
```
























<!-- SECTION: Module 7 (M7) -->





### **Module 7 (M7) - Spotify II**




**How does this tie to the journey?** 


*Purpose: What is Statistics/Data Science (DS) Question we want to answer?*

3. **Danceability and Energy relationship**: Is there a relationship between danceability and energy across all tracks using only Billie Ellish and Dua Lipa Songs?

- *Manipulation*: Select the appropriate variables and filter based on the selected artists, Billie Ellish and Dua Lipa? 

- *Visualization*: Use a scatter plot to visualize the relation between danceability and energy of Billie Ellish and Dua Lipa songs.



##### The Extraction
```{r}
library(gato365dsh2024)
library(dplyr)
library(ggplot2)
```


##### The Tranformation
```{r}
dua_lipa_billie_eilish_energy_danceability_1_df <- all_artists_df %>% 
  select(artist_name, energy,danceability) %>% 
  filter(artist_name %in% c("Dua Lipa","Billie Eilish")) 
```

##### The Visualization
```{r}
dua_lipa_billie_eilish_energy_danceability_1_df %>% 
  ggplot(aes(x = energy, y = danceability, color = artist_name)) +
  geom_point() +
  geom_vline(xintercept = 0.5) +
  geom_hline(yintercept = 0.5) +
  xlim(0,1) +
  ylim(0,1)+
  theme_custom()
```





<!-- SECTION: Module 8 (M8) -->














#### ----------------------------------


### **Module 8 (M8) - Sports II**

*Purpose: What is Statistics/Data Science (DS) Question we want to answer?*


3. **Touchdown to Interception Ratio**: Which quarterbacks have a touchdown to interception ratio greater than 2?
- *Manipulation*: Calculate the touchdown to interception ratio for each quarterback and filter those with a ratio greater than 2.
- *Visualization*: Create a scatter plot with touchdowns on the x-axis and interceptions on the y-axis, highlighting quarterbacks who meet the criteria.




##### The Extraction

```{r}
library(tidyverse) ## ggplot2 and dplyr
library(gato365dsh2024)
```

##### The Tranformation

```{r}
td_per_int_passing_df <- passing_stats_df %>% 
  distinct(.,.keep_all = TRUE) %>% 
  filter(year == 2023) %>% 
  select(player,int,td)  %>% 
  mutate(td_per_int = ifelse(int == 0, NA, td/int)) %>% 
  filter(td_per_int <=4 & td_per_int >=1) %>% 
  arrange(td_per_int) %>% 
  mutate(player = factor(player, levels = unique(player)))
```

##### The Visualization


```{r}
td_per_int_passing_df  %>% 
  ggplot(aes(x=player, y = td_per_int)) +
  geom_point() + 
  geom_text(aes(label=round(td_per_int, 2)), hjust=-0.1, vjust=-0.5) +
  labs(x = "Player", y = "TD/INT Ratio", title = "TD/INT Ratio for Each Player in 2023") +
  coord_flip() +
  theme_custom()
```















## Weekend 3 Outline


More questions



##### The Extraction

##### The Tranformation

##### The Visualization


- Hackathon




