---
title: "Project Title Here"
author: "Author Name Here"
date: "Current Date Here"
---






### **Module 7 (M7) - Spotify II**




**How does this tie to the journey?** 


*Purpose: What is Statistics/Data Science (DS) Question we want to answer?*

3. **Danceability and Energy relationship**: Is there a relationship between danceability and energy across all tracks using only Billie Ellish and Dua Lipa Songs?

- *Manipulation*: Select the appropriate variables and filter based on the selected artists, Billie Ellish and Dua Lipa? 

- *Visualization*: Use a scatter plot to visualize the relation between danceability and energy of Billie Ellish and Dua Lipa songs.



##### The Extraction
```{r}
library(gato365dsh2024)
library(dplyr)
library(ggplot2)
```


##### The Tranformation
```{r}
dua_lipa_billie_eilish_energy_danceability_1_df <- all_artists_df %>% 
  select(artist_name, energy,danceability) %>% 
  filter(artist_name %in% c("Dua Lipa","Billie Eilish")) 
```

##### The Visualization
```{r}
dua_lipa_billie_eilish_energy_danceability_1_df %>% 
  ggplot(aes(x = energy, y = danceability, color = artist_name)) +
  geom_point() +
  geom_vline(xintercept = 0.5) +
  geom_hline(yintercept = 0.5) +
  xlim(0,1) +
  ylim(0,1)+
  theme_custom()
```





