---
title: "Project Title Here"
author: "Author Name Here"
date: "Current Date Here"
---



## Weekend 2 Outline





### **Module 5 (M5) - Spotify I**


#### **Part 1 (P1)** ETV - Spotify I
**How does this tie to the journey?** As a Spotify data engineer, your role is pivotal in understanding the current trends of popular artists. By analyzing data such as user listening habits, artist popularity metrics, and song play counts, you can identify patterns and trends among these artists. These insights not only help Spotify curate better playlists and make more accurate song recommendations, but they also contribute to predicting future hits. Thus, your work directly influences the user experience on Spotify, making it more personalized and enjoyable for listeners worldwide. This is how your task ties into your journey as a data engineer at Spotify.

*Purpose: What is Statistics/Data Science (DS) Question we want to answer?*
1. **Artist Album Release Patterns**: How has the number of tracks on albums released by each artist ("Dua Lipa" and "Olivia Rodrigo") changed over the years?
- *Data Manipulation*: Filter the data for the artists "Dua Lipa" and "Olivia Rodrigo". Remove duplicate tracks based on track name, then group the data by artist and album name. Count the number of tracks on each album and record the album release year.
- *Visualization*: Create a line graph showing the number of tracks on albums released per year for each artist. The lines are colored differently for each artist.


##### The Extraction
```{r}
library(gato365dsh2024)
library(dplyr)
glimpse(all_artists_df)
unique(all_artists_df$artist_name)
```

##### The Transformation
```{r}
library(ggplot2)
artist_album_df <- all_artists_df %>% 
  filter(artist_name %in% c("Dua Lipa","Olivia Rodrigo")) %>% 
  distinct(track_name, .keep_all = TRUE) %>%  # Remove duplicates based on artist and album
  group_by(artist_name, album_name) %>%  # Group by artist and album year
  summarise(count = n(), album_release_year = first(album_release_year))  # Count the number of albums
```


##### The Visualization 
```{r}
artist_album_df %>% 
  ggplot(aes(x = album_release_year, y = count, group = artist_name, color = artist_name)) +  # Plot album_year on x-axis and count on y-axis, group and color by artist
  geom_line() +  # Use line graph
  geom_point() +
  scale_color_manual(values = c("blue","green")) +
  labs(title = "Artist Album Release Patterns Over the Years", x = "Year", y = "Number of Tracks on Albums", color = "Artist Name") +  # Add labels
  theme_custom()
```














