---
title: "Project Title Here"
author: "Author Name Here"
date: "Current Date Here"
---





### **Module 6 (M6) - Sports I**

*Purpose: What is Statistics/Data Science (DS) Question we want to answer?*

#### **Part 1 (P1)**: Understanding Age Distribution


**How does this tie to the journey?** 

As a data analyst for the NBA league, your role is crucial in understanding the age distribution of players. By analyzing player data, you can identify patterns and trends among different age groups. This insight not only helps in understanding the career longevity and prime playing years of players, but also aids in strategic team management and talent scouting. Thus, your work directly influences the strategic decisions made in the NBA league, contributing to the success of teams and the overall competitiveness of the league. This is how your task ties into your journey as a data analyst at the NBA.

*Purpose: What is Statistics/Data Science (DS) Question we want to answer?*

2a. **Age Distribution of Players**: How does the age distribution look across all players, and which age group is most prevalent? 

- *Data Manipulation*:  Group the players by age and count the number of players in each age group.

- *Visualization*:  Create a histogram to display the age distribution.
















##### The Extraction

```{r}
library(gato365dsh2024)
library(dplyr)
library(ggplot2)
```


##### The Transformation

```{r}
nba_players_age_df <- nba_year_2023_df %>% 
  distinct(player_name, .keep_all = TRUE) %>%  # Remove duplicates based on the player with the most games
  mutate(age_group = cut(age, breaks = seq(15, 50, 5), include.lowest = TRUE)) %>%  # Create arbitrary age groups
  group_by(age_group) %>%  # Group by age group
  summarise(freq = n()) 
```
##### The Visualization

```{r}
nba_players_age_df %>% 
  ggplot(aes(x = age_group, y = freq)) +  # Plot age_group on x-axis and frequency on y-axis
  geom_bar(stat = "identity", fill = "steelblue") +  # Use bar chart
  labs(title = "Age Distribution of NBA Players in 2023", x = "Age Group", y = "Number of Players") +  # Add labels
  theme_custom()

```




#### ----------------------------------

#### **Part 2 (P2)**: Age Distribution by Position


**How does this tie to the journey?**  
Examining age distribution by player position uncovers patterns of career duration and potential position-specific physical demands, guiding strategic decisions in team composition and player development.



```{r}
## 1b. **Age and Position Distribution of Players**: How does the age distribution look across all players conditioned on position?
nba_year_2023_df %>% 
  distinct(player_name, .keep_all = TRUE) %>%  # Remove duplicates based on the player with the most games
  mutate(position = str_split(position, "-", simplify = TRUE)[,1]) %>% 
  mutate(age_group = cut(age, breaks = seq(15, 50, 5), include.lowest = TRUE)) %>%  # Create arbitrary age groups
  group_by(position, age_group) %>%  # Group by position and age group
  summarise(freq = n()) %>%  # Count the frequency in each group
  ggplot(aes(x = age_group, y = freq, fill = position)) +  # Plot age_group on x-axis and frequency on y-axis, fill by position
  geom_bar(stat = "identity", position = "dodge") +  # Use bar chart with bars side-by-side
  labs(title = "Age Distribution of NBA Players in 2023 by Position", x = "Age Group", y = "Number of Players") +  # Add labels
  theme_custom()
```
























